The \emph{FrobeniusThresholds} package for \emph{Macaulay2}

### 2019-11-22 Fsing/FrobeniusThresholdsPaper/FrobeniusThresholds.tex
Updated links to our packages to be permalinks

This paper describes the \emph{Macaulay2} package \emph{FrobeniusThresholds} \cite{M2, FThresholdsPackage}, which provides tools for computing or estimating certain fundamental invariants in positive characteristic commutative algebra, namely \emph{$F$-pure thresholds}, \emph{$F$-thresholds}, and \emph{$F$-jumping exponents}.
Recall that a ring of prime characteristic $p>0$ is \emph{$F$-pure} if the Frobenius map|that is, the ring endomorphism sending an element to its $p$-th power|is a pure morphism;  under natural geometric hypotheses, this is equivalent to the condition that the Frobenius morphism splits as a map of rings.
The concept of $F$-purity has proven to be important in commutative algebra, and has a rich history.
Indeed, it first appeared in the work of Hochster and Roberts to study local cohomology \cite{HochsterRobertsFrobeniusLocalCohomology}, was compared with rational singularities in  \cite{FedderFPureRat}, and was used to study global properties of Schubert varieties in \cite{MehtaRamanathanFrobeniusSplittingAndCohomologyVanishing}.

@@

It turns out that the sequence $\big(\nu_f^{\idealm}(p^e) \big)_{e=0}^{\infty}$ itself, and not just its limit, encodes interesting information about $f$.  For example, it is closely related to the \emph{Bernstein--Sato polynomial} of $f$ \cite{MustataTakagiWatanabeFThresholdsAndBernsteinSato}.  Thus, it is natural to ask whether one can recover the sequence $\big( \nu_f^{\idealm}(p^e) \big)_{e=0}^{\infty}$ from the limit $\ft^{\idealm}(f)$.  Fortunately, the answer to this question is yes \cite{MustataTakagiWatanabeFThresholdsAndBernsteinSato, HernandezFPurityOfHypersurfaces}; for each $e$ we have
\begin{equation*}
\nu_f^{\idealm}(p^e) = \lceil p^e \cdot \ft^{\idealm}(f) \rceil - 1.
\end{equation*}

@@

\subsection*{Acknowledgements.}  The authors enthusiastically thank everyone who helped complete the \emph{FrobeniusThresholds} package: the package coauthors Juliette Bruce and Daniel Smolkin, and contributors  Erin Bela, Zhibek Kadyrsizova, Moty Katzman, Sara Malec, and Marcus Robinson.

@@

We first consider the \texttt{nu} function, a fundamental component of the package \emph{FrobeniusThresholds}.
We adopt the setup established in the introduction: we work in a polynomial ring $R$ over a finite field of characteristic $p>0$,  $\idealm$ denotes the ideal generated by the variables, and $e$ is a natural number.
If $I$ and $J$ are ideals of $R$, the command \texttt{nu(e,I,J)} outputs the extended integer $\nu_I^J(p^e)$ defined in the introduction; if $f$ is an element of $R$, \texttt{nu(e,f,J)} outputs $\nu_f^J(p^e) \coloneqq \nu_{\langle f \rangle}^J(p^e)$.
When the third argument is omitted from the function \texttt{nu}, it is assumed to be the maximal ideal $\idealm$.

>>>>

This paper describes the \emph{Macaulay2} package \emph{FrobeniusThresholds}\footnote{This package was previously named \emph{FThresholds}} \cite{M2, FThresholdsPackage}, which provides tools for computing or estimating certain fundamental invariants in positive characteristic commutative algebra, namely \emph{$F$-pure thresholds}, \emph{$F$-thresholds}, and \emph{$F$-jumping exponents}.
Recall that a ring of prime characteristic $p>0$ is \emph{$F$-pure} if the Frobenius map|that is, the ring endomorphism sending an element to its $p$-th power|is a pure morphism;  under natural geometric hypotheses, this is equivalent to the condition that the Frobenius morphism splits as a map of rings.
The concept of $F$-purity has proven to be important in commutative algebra, and has a rich history.
Indeed, it first appeared in the work of Hochster and Roberts to study local cohomology \cite{HochsterRobertsFrobeniusLocalCohomology}, was compared with rational singularities in  \cite{FedderFPureRat}, and was used to study global properties of Schubert varieties in \cite{MehtaRamanathanFrobeniusSplittingAndCohomologyVanishing}.

@@

It turns out that the sequence $\big(\nu_f^{\idealm}(p^e) \big)_{e=0}^{\infty}$ itself, and not just its limit, encodes interesting information about $f$.  For example, it is closely related to the \emph{Bernstein--Sato polynomial} of $f$ \cite{MustataTakagiWatanabeFThresholdsAndBernsteinSato}.  
Remarkably, 
one can recover the sequence $\big( \nu_f^{\idealm}(p^e) \big)_{e=0}^{\infty}$ from the limit $\ft^{\idealm}(f)$ \cite{MustataTakagiWatanabeFThresholdsAndBernsteinSato, HernandezFPurityOfHypersurfaces}: For each $e$, we have
\begin{equation*}
\nu_f^{\idealm}(p^e) = \lceil p^e \cdot \ft^{\idealm}(f) \rceil - 1.
\end{equation*}

@@

\subsection*{Acknowledgements.}  The authors enthusiastically thank everyone who helped complete the \emph{FrobeniusThresholds} package: the package coauthors Juliette Bruce and Daniel Smolkin, and contributors  Erin Bela, Zhibek Kadyrsizova, Moty Katzman, Sara Malec, and Marcus Robinson.  We also thank the authors of the \emph{TestIdeals} package, which, beyond the authors of the present paper and those listed above, are Alberto Boix, Drew Ellingson, Matthew Mastroeni, and Maral Mostafazadehfard.

@@

We first describe the \texttt{nu} function, a fundamental component of the package \emph{FrobeniusThresholds}.
We adopt the setup established in the introduction: we work in a polynomial ring $R$ over a finite field of characteristic $p>0$,  $\idealm$ denotes the ideal generated by the variables, and $e$ is a natural number.
If $I$ and $J$ are ideals of $R$, the command \texttt{nu(e,I,J)} outputs the extended integer $\nu_I^J(p^e)$ defined in the introduction; if $f$ is an element of $R$, \texttt{nu(e,f,J)} outputs $\nu_f^J(p^e) \coloneqq \nu_{\langle f \rangle}^J(p^e)$.
When the third argument is omitted from the function \texttt{nu}, it is assumed to be the maximal ideal $\idealm$.
